---
title: "Single-Index Conditional Distribution Model Fitting"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{demoR_SICDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>")
```

```{=tex}
\newcommand{\E}{\mathrm{E}}
\newcommand{\pr}{\mathrm{pr}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\cond}{\,|\,}
\newcommand{\T}{\intercal}
\newcommand{\diff}{\mathrm{d}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bfxy}[1]{\mathbf{#1}}
\newcommand{\bfX}{\mathbf{X}}
\newcommand{\bfx}{\mathbf{x}}
```

# Package

```{r setup}
library(MYHRcpp)
```

# Data Generating

1. $X\sim N_p(0,1)$.
2. $Y=\sin(X^\T\beta_0)+\varepsilon$, where $\beta_0=(1,\dots,1)$ and $\varepsilon\sim N(0, 0.2)$.

```{r}
n <- 100
p <- 5

X <- matrix(rnorm(n*p), nrow = n, ncol = p)
Y <- as.matrix(sin(X %*% rep(1, p))+rnorm(n, mean = 0, sd = 0.2))
```

# Cumulative Sliced Inverse Regression

This is the method proposed in the following paper:

- Zhu, Zhu, and Feng (2010). Dimension Reduction in Regressions Through Cumulative Slicing Estimation. \textit{Journal of the American Statistical Association}. \textbf{105}, 1455--1466.

The first direction in the resulting basis estimates the coefficients of the single-index.

```{r}
SDR1 <- cumuSIR(X = X, Y = Y)
SDR1
```

# Single-Index Conditional Distribution Model Fitting

Options:

- `initial`: Initial vector for $\beta_0$, which is of the same dimension as $X$
- `kernel`: `"K2_Biweight"` or `"K4_Biweight"`

```{r}
SID1 <- SIDR(X = X, Y = Y)
SID2 <- SIDR(X = X, Y = Y, initial = SDR1$basis[, 1])
SID3 <- SIDR(X = X, Y = Y, initial = rep(1, p))
SID4 <- SIDR(X = X, Y = Y, initial = rep(1, p), kernel = "K2_Biweight")
SID5 <- SIDR(X = X, Y = Y, initial = rep(1, p), bandwidth = 1)
```

Calculation of the estimated conditional distribution function

```{r}
Fhat <- NW(X = X %*% SID1$coef, Y = Y, x = X %*% SID1$coef,
           regression = "distribution", y = sort(Y),
           bandwidth = SID1$bandwidth)
```

