---
title: "Maximum Likelihood Estimation of Generalized Linear Models"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{MLE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>")
```

```{=tex}
\newcommand{\E}{\mathrm{E}}
\newcommand{\pr}{\mathrm{pr}}
\newcommand{\var}{\mathrm{var}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\cond}{\,|\,}
\newcommand{\T}{\intercal}
\newcommand{\diff}{\mathrm{d}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bfxy}[1]{\mathbf{#1}}
\newcommand{\bfX}{\mathbf{X}}
\newcommand{\bfx}{\mathbf{x}}
```

# Package

```{r}
library(MYHRcpp)
```

# Notations

1.  $Y\in\mathbb{R}$ is the response of interest.
2.  $\bfX\in\mathbb{R}^p$ is the covariate vector.

# Normal regression model

The normal regression model assumes that $Y$ conditional on $\bfX=\bfx$ follows a normal distribution with mean $\alpha_0+\bfx^\T\beta_0$ and variance $\sigma_0^2$.

## Simulated data

The function `simGLM.normal()` is used to generate simulated data from the normal regression model.

```{r}
n <- 50
p <- 3
alpha0 <- 1
beta0 <- rep(0.2, p)
sigma0 <- 0.1

set.seed(321)
X <- rnorm(n*p, mean = 0, sd = 1)
dim(X) <- c(n, p)
data.test <- simGLM.normal(Xi = X,
                           alpha0 = alpha0,
                           beta0 = beta0,
                           sigma0 = sigma0)
data.test[1:5, ]
data.test <- simGLM.normal(Xi = X,
                           alpha0 = alpha0,
                           beta0 = beta0,
                           sigma0 = sigma0,
                           seed = 123)
data.test[1:5, ]
```

## Estimation

The function `MLE.normal()` is used to find the maximum likelihood estimators for $\alpha_0$, $\beta_0$, and $\sigma_0$.

```{r}
test1.MLE <- MLE.normal(X = data.test[paste("covariate", 1:p, sep=".")],
                        Y = data.test["response"])
test1.MLE

test2.MLE <- MLE.normal(X = data.test[paste("covariate", 1:p, sep=".")],
                        Y = data.test["response"],
                        initial = c(-1, rep(1, p), 1),
                        wi.boot = rep(1, n),
                        do.SE = TRUE,
                        X.future = matrix(0, nrow = 1, ncol = p))
test2.MLE

# simulations

SN <- 10
BN <- 10

test.MLE.MC <- array(0, c(p+2, 3, SN))
dimnames(test.MLE.MC) <- list(c("alpha", paste("beta", 1:p, sep = ""), "sigma"),
                              c("coefficient", "SE", "BSE"),
                              NULL)

for (sn in 1:SN)
{
  set.seed(531+sn)

  X <- rnorm(n*p, mean = 0, sd = 1)
  dim(X) <- c(n, p)
  data.test <- simGLM.normal(Xi = X,
                             alpha0 = alpha0,
                             beta0 = beta0,
                             sigma0 = sigma0,
                             seed = 123+sn)

  test.MLE <- MLE.normal(X = data.test[paste("covariate", 1:p, sep=".")],
                         Y = data.test["response"],
                         do.SE = TRUE)
  test.MLE.MC[, "coefficient", sn] <- test.MLE$parameter
  test.MLE.MC[, "SE", sn] <- diag(test.MLE$Cov.coef)^0.5

  test.MLE.boot <- array(0, c(p+2, BN))
  for (bn in 1:BN)
  {
    set.seed(sn+bn+123)

    rw <- rexp(n, rate = 1)
    rw <- n*rw/sum(rw)

    test.MLE.boot[, bn] <-
      MLE.normal(X = data.test[paste("covariate", 1:p, sep=".")],
                 Y = data.test["response"],
                 wi.boot = rw)$parameter

    #print(paste("simulation", sn, "boot", bn, sep = ""))
  }
  test.MLE.MC[, "BSE", sn] <- apply(test.MLE.boot, 1, sd)

  #print(sn)
}

test.MLE.MC.summary <- array(0, c(p+2, 4))
dimnames(test.MLE.MC.summary) <-
  list(c("alpha", paste("beta", 1:p, sep = ""), "sigma"),
       c("coefficient", "SD", "SE", "BSE"))
test.MLE.MC.summary[, c("coefficient", "SE", "BSE")] <-
  apply(test.MLE.MC, c(1, 2), mean)
test.MLE.MC.summary[, "SD"] <-
  apply(test.MLE.MC, c(1, 2), sd)[, "coefficient"]
test.MLE.MC.summary
```

# Logistic regression model

The logistic regression model assume that $Y\in\{0,1\}$ conditional on $\bfX=\bfx$ follows a Bernoulli distribution with $$\pr(Y=1)=\frac{\exp(\alpha_0+\bfx^\T\beta_0)}{1+\exp(\alpha_0+\bfx^\T\beta_0)}.$$

## Simulated data

The function `simGLM.logistic()` is used to generate simulated data from the logistic regression model.

```{r}
n <- 50
p <- 3
alpha0 <- 0.1
beta0 <- rep(0.2, p)

set.seed(321)
X <- rnorm(n*p, mean = 0, sd = 1)
dim(X) <- c(n, p)
data.test <- simGLM.logistic(X = X,
                             alpha0 = alpha0,
                             beta0 = beta0)
data.test[1:5, ]
data.test <- simGLM.logistic(X = X,
                             alpha0 = alpha0,
                             beta0 = beta0,
                             seed = 123)
data.test[1:5, ]
```

## Estimation

The function `MLE.logistic()` is used to find the maximum likelihood estimators for $\alpha_0$ and $\beta_0$.

```{r}
test1.MLE <- MLE.logistic(X = data.test[paste("covariate", 1:p, sep=".")],
                          Y = data.test["response"])
test1.MLE

test2.MLE <- MLE.logistic(X = data.test[paste("covariate", 1:p, sep=".")],
                          Y = data.test["response"],
                          initial = c(0.1, rep(0.2, p)),
                          do.SE = TRUE,
                          X.future = matrix(0, nrow = 1, ncol = p))
test2.MLE

# simulations

SN <- 10
BN <- 10

test.MLE.MC <- array(0, c(p+1, 3, SN))
dimnames(test.MLE.MC) <- list(c("alpha", paste("beta", 1:p, sep = "")),
                              c("coefficient", "SE", "BSE"),
                              NULL)

for (sn in 1:SN)
{
  set.seed(531+sn)

  X <- rnorm(n*p, mean = 0, sd = 1)
  dim(X) <- c(n, p)
  data.test <- simGLM.logistic(Xi = X,
                               alpha0 = alpha0,
                               beta0 = beta0,
                               seed = 123+sn)

  test.MLE <- MLE.logistic(X = data.test[paste("covariate", 1:p, sep=".")],
                           Y = data.test["response"],
                           do.SE = TRUE, diff.tol = 1e-3)
  test.MLE.MC[, "coefficient", sn] <- test.MLE$parameter
  test.MLE.MC[, "SE", sn] <- diag(test.MLE$Cov.coef)^0.5

  test.MLE.boot <- array(0, c(p+1, BN))
  for (bn in 1:BN)
  {
    set.seed(sn+bn+123)

    rw <- rexp(n, rate = 1)
    rw <- n*rw/sum(rw)

    test.MLE.boot[, bn] <-
      MLE.logistic(X = data.test[paste("covariate", 1:p, sep=".")],
                   Y = data.test["response"],
                   wi.boot = rw)$parameter

    #print(paste("simulation", sn, "boot", bn, sep = ""))
  }
  test.MLE.MC[, "BSE", sn] <- apply(test.MLE.boot, 1, sd)

  #print(sn)
}

test.MLE.MC.summary <- array(0, c(p+1, 4))
dimnames(test.MLE.MC.summary) <-
  list(c("alpha", paste("beta", 1:p, sep = "")),
       c("coefficient", "SD", "SE", "BSE"))
test.MLE.MC.summary[, c("coefficient", "SE", "BSE")] <-
  apply(test.MLE.MC, c(1, 2), mean)
test.MLE.MC.summary[, "SD"] <-
  apply(test.MLE.MC, c(1, 2), sd)[, "coefficient"]
test.MLE.MC.summary
```
